---
title             : |
  (Im)polite Requests in L2 Spanish
shorttitle        : "L2 Spanish Requests"
author            : 
  - name: Robert Esposito
    corresponding: true
    affiliations: Rutgers University
    email: rme70@rutgers.edu

bibliography      : "./lit/polite_bib.bib"
csl               : "./lit/apa-6th-edition.csl"
indent: true

format:
  docx:
    reference-doc: "./ref_doc/apa_style.docx"
    number-sections: false
    highlight-style: github
    code-line-numbers: true
    fig-align: center
    tbl-colwidths: true
    mermaid:
      theme: neutral
    mermaid-format: png

editor_options: 
  markdown: 
    wrap: sentence
metadata: 
  subject: linguistics
  description: "L2 Intonational Phonology"
  category: article
---

```{r}
#| label: setup
#| include: false

# Set document defaults
knitr::opts_chunk$set(
  cache.extra = knitr::rand_seed, 
  fig.retina = 2, 
  echo = F, 
  message = F, 
  warning = F,
  fig.asp = 0.5625,
  out.width = "100%",
  dpi = 300, 
  fig.path = 'figs/', 
  dev = c("png", "pdf")
  )

library("here")
library("fs")
library("dplyr")
library("tidyr")
library("stringr")
```

{{< pagebreak >}}

# Author note {.unnumbered}

Correspondence concerning this article should be addressed to Robert Esposito, Rutgers University - Department of Spanish and Portuguese, 15 Seminary Place, New Brunswick, NJ 08904, USA . E-mail: rme70@rutgers.edu.

{{< pagebreak >}}

# Abstract {.unnumbered}

This is where the abstract will go.

*Keywords:* Pitch, Prosody, Intonation, Intonational Phonology, Pragmatics, Linguistics

*Word count:* WORD_COUNT

{{< pagebreak >}}


# `r rmarkdown::metadata$title`

# Introduction

F0, and its phonological correlate pitch, is used cross-linguistically for a variety of functions. Typologically, some languages, like Yoruba or Mandarin Chinese, map pitch to tones, which function like segmental phonemes, used to deliver semantic or morphosyntactic meaning. In other languages, like the vast majority of Indo-European languages, pitch is used for a variety of pragmatic purposes, in which case it is labeled as "intonation". For example, intonation can be used to differentiate sentence modality or expressing doubt [@ladd2008intonational]. Another use, under investigation here, is to express politeness [@brown1987politeness].

Intonation in modern linguistics is only recently receiving more attention. Where previous models treated intonation with the goals of speech synthesis [@hart1990perceptual] or pedagogy [@pike1945intonation], @ladd2008intonational and colleagues spearheaded the field of intonational phonology and developed the Autosegmental Metrical theory of intonational phonology (henceforth AM) from which to study intonation from a phonological perspective.

The vast majority of research on intonation investigate monolingual speakers, but there do exist recent works that investigate L2 intonation strategies [@astruc2016intonational], as well as attempt to model L2 intonation [@mennen2015beyond]. The current research undertaken here attempts to investigate the L2 intonation system by looking at L1 English L2 Spanish intonational strategies to express varying levels of politeness.

## Autosegmental Metrical Theory

Within the AM framework [@ladd2008intonational], the continuous phonetic cue F0 is mapped to ordered strings of Low (L) and High (H) tones that attach to stressed syllables, called pitch accents, and edges of phrases, called boundary tones. Significantly, the phonological representations are not meant to include all elements of pitch modulation, just as all phonetic elements are not recorded for segments. That is, phonetically distinct contours may be realization of the very same underlying phonological element. Thus, the goal of AM is not to record every phonetic detail of F0, but to understand how F0 maps to discrete, categorical tones that express meaning [@ladd2022trouble].

The abstract, phonological representation of intonation with the AM framework is carried out through the Tones and Breaks Indices (ToBI) tool, originally developed for American English (@silverman1992tobi). This tool has also been developed for Spanish, called Sp_ToBI (@beckman2002intonation). Of note is that ToBI requires that each variety of a language have its own ToBI. For example, various varieties of English [e.g., @mayo1997prosodic, @silverman1992tobi] and Spanish (@prieto2010transcription) have since been documented using ToBI.

@fig-tobi-example gives an example of an utterance labeled using ToBI. ToBI associates a sound file with its spectrogram and pitch track, as well as several tiers of intonation. The typical structure is a transcription tier, typically divided into syllables and represented with either the language's orthography or IPA, a tone tier, where pitch accents and boundary tones are aligned with their associated syllables, and optionally a break index tier, representing the perceived strength of prosodic boundaries. 

```{r}
#| label: fig-tobi-example
#| fig-cap: The utterance "Ana lleva el abrigo" ("Ana is wearing a coat") labeled using ToBI.
knitr::include_graphics(
  here("figs", "tobi_example.png")
  )
```

On the tone tier, pitch accents are represented with * (e.g. H\* represents a High tone associated with a stressed syllable), while % is associated with boundary tones (e.g. H% is a High boundary tone at the end of a phrase). A - is used to represent a phrase boundary that is not utterance final (e.g. H- is a High phrase boundary). As mentioned above, tones can be combined to represent rises and falls, and when used as pitch accents, the * indicates the tone that is most associated with the stressed syllable. For example, H\*L is a fall, where the peak is within the stressed syllable; HL\* represents a fall where the peak occurs either before or at the very start of the stressed syllable.

## Pitch and Politeness

Although AM takes for granted the arbitrary nature of sound to meaning correspondences that dominate human language, some researchers propose that some aspects of pitch-meaning correspondences are non-arbitrary. For example, @ohala1983cross, in his "frequency code", the use of pitch in human language evolved from a cross-species use of F0, where higher pitch is "nonthreatening, submissive, subordinate, in need of the receiver's cooperation and good will", whereas lower pitch is "aggressive, assertive..." (p. 8). He claims that these correspondences are cross-linguistic, citing the universality of F0's use to distinguish sentence modality. That is, cross-linguistically, low pitch is associated with statements ("assertiveness"), whereas high pitch is associated with questions ("in need of the receiver's cooperation and good will"). Under this belief, using higher pitch is a cross-linguistic method of expressing politeness.

This belief is supported by @brown1987politeness. They give Tzeltal as an example, where politeness is marked by high pitch or falsetto. They claim elsewhere that there is a universal association between high pitch and tentativeness [@brown1974theory].

Given the cited research, it would be expected that some of the earliest politeness strategies adopted by L2 learners would involve pitch, as opposed to (morpho-)syntax or semantics, as they already have the correspondence between higher pitch and politeness available to them in their L1. This prediction will be further developed below in the discussion of politeness strategies and L1 to L2 intonation transfer.

## Politeness

@brown1974theory provides a thorough, cross-linguistic investigation of politeness.

## LiLT model

## Previous Research

# Scraps


Given these proposals, it would be expected that prosodic features may be some of the first politeness strategies positively transferred from the L1 to the L2, due to the crosslinguistic similarities and ready availability. A variety of politeness strategies require more complex morpho-syntactic transformations (cite brown here), but something as simple as using higher pitch to express increasing politeness is something available to even the most neophyte language learner. This study aims to investigate the intonational resources that L2 Spanish speakers have available to them.



{{< pagebreak >}}


# References {.unnumbered}

::: {#refs}
:::